### 2018-09-21 manuscript/manuscript_template.Rmd
Update markdown manuscript template

<!DOCTYPE html>
<html lang="en">
  <head>
    <meta charset="utf-8">
  <link rel="dns-prefetch" href="https://assets-cdn.github.com">
  <link rel="dns-prefetch" href="https://avatars0.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars1.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars2.githubusercontent.com">
  <link rel="dns-prefetch" href="https://avatars3.githubusercontent.com">
  <link rel="dns-prefetch" href="https://github-cloud.s3.amazonaws.com">
  <link rel="dns-prefetch" href="https://user-images.githubusercontent.com/">

  <link crossorigin="anonymous" media="all" integrity="sha512-/yZA7MGD9acBi5MmgjnivwjGeiNsmimsI/bhLlK7akrRuw+hMxNpJoQEMO9Zt0Nq6ZRu1m1NGV6r95Ib0zx7Qw==" rel="stylesheet" href="https://assets-cdn.github.com/assets/frameworks-1ca00d32d1a8adc78ae7bb6677410eb1.css" />
  <link crossorigin="anonymous" media="all" integrity="sha512-pHF1luoMSUFxvz63bW3jEbbifYl/pruHUw+kljThAPi2g0nhgMdvWj4bQRh4O8MYp9NR60XYcylaLXmYA8YXSw==" rel="stylesheet" href="https://assets-cdn.github.com/assets/github-5043b3d96d50b119b971c2cdde4e0a62.css" />

  <meta name="viewport" content="width=device-width">

  <title>Blog/manuscript_template.Rmd at master · petrkeil/Blog</title>
    <meta name="description" content="This repository contains codes for all of my blog posts since March 11 2015. - petrkeil/Blog">
    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
  <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
  <meta property="fb:app_id" content="1401488693436528">

    <meta property="og:image" content="https://avatars2.githubusercontent.com/u/5421227?s=400&amp;v=4" /><meta property="og:site_name" content="GitHub" /><meta property="og:type" content="object" /><meta property="og:title" content="petrkeil/Blog" /><meta property="og:url" content="https://github.com/petrkeil/Blog" /><meta property="og:description" content="This repository contains codes for all of my blog posts since March 11 2015. - petrkeil/Blog" />

  <link rel="assets" href="https://assets-cdn.github.com/">
  <link rel="web-socket" href="wss://live.github.com/_sockets/VjI6MzE5MjkzMDMyOmY4ZDEyYmU0MjE5YTNjMTg5YzMwODMwOTRhOWJmYjA2NmU1ZDRjYzNmNTlkMmZmNThiNmEyMDhkY2U0ODM3YjU=--6699b174fe2e40623d9cdaf2af3c8eb14a2753a8">
  <meta name="pjax-timeout" content="1000">
  <link rel="sudo-modal" href="/sessions/sudo_modal">
  <meta name="request-id" content="DEC4:7B80:13DB109:2408B1B:5BA52F38" data-pjax-transient>

  <meta name="selected-link" value="repo_source" data-pjax-transient>

      <meta name="google-site-verification" content="KT5gs8h0wvaagLKAVWq8bbeNwnZZK1r1XQysX3xurLU">
    <meta name="google-site-verification" content="ZzhVyEFwb7w3e0-uOTltm8Jsck2F5StVihD0exw2fsA">
    <meta name="google-site-verification" content="GXs5KoUUkNCoaAZn7wPN-t01Pywp9M3sEjnt_3_ZWPc">

  <meta name="octolytics-host" content="collector.githubapp.com" /><meta name="octolytics-app-id" content="github" /><meta name="octolytics-event-url" content="https://collector.githubapp.com/github-external/browser_event" /><meta name="octolytics-dimension-request_id" content="DEC4:7B80:13DB109:2408B1B:5BA52F38" /><meta name="octolytics-dimension-region_edge" content="iad" /><meta name="octolytics-dimension-region_render" content="iad" /><meta name="octolytics-actor-id" content="34245401" /><meta name="octolytics-actor-login" content="jamesgrecian" /><meta name="octolytics-actor-hash" content="821728246d06cdfe784e56dfb8dc2a4447ec5ee2fb7525c8728616156ffa29a2" />
<meta name="analytics-location" content="/&lt;user-name&gt;/&lt;repo-name&gt;/blob/show" data-pjax-transient="true" />

    <meta name="google-analytics" content="UA-3769691-2">

  <meta class="js-ga-set" name="userId" content="078bbd73f63945865b3a6f9e3c674706" %>

<meta class="js-ga-set" name="dimension1" content="Logged In">

      <meta name="hostname" content="github.com">
    <meta name="user-login" content="jamesgrecian">

      <meta name="expected-hostname" content="github.com">
    <meta name="js-proxy-site-detection-payload" content="NGYyMDUzNWVkYjVhZGU3ZTI0NmJjNzE5MmZkYTM3N2YwMDRlOTk2MTFmMDRkYjNjY2VjZDJlYTIzN2JlYTFlN3x7InJlbW90ZV9hZGRyZXNzIjoiMTU2LjM0LjI0OC4yMDciLCJyZXF1ZXN0X2lkIjoiREVDNDo3QjgwOjEzREIxMDk6MjQwOEIxQjo1QkE1MkYzOCIsInRpbWVzdGFtcCI6MTUzNzU1MjE5NiwiaG9zdCI6ImdpdGh1Yi5jb20ifQ==">

    <meta name="enabled-features" content="DASHBOARD_V2_LAYOUT_OPT_IN,EXPLORE_DISCOVER_REPOSITORIES,UNIVERSE_BANNER,MARKETPLACE_PLAN_RESTRICTION_EDITOR,MARKETPLACE_RETARGETING,QUOTE_MARKDOWN">

  <meta name="html-safe-nonce" content="4d33eb369dbc9e939c0c70269799ff90a62c9762">

  <meta http-equiv="x-pjax-version" content="3842ef7af97baac02526bd1bb609fd99">

      <link href="https://github.com/petrkeil/Blog/commits/master.atom" rel="alternate" title="Recent Commits to Blog:master" type="application/atom+xml">

  <meta name="go-import" content="github.com/petrkeil/Blog git https://github.com/petrkeil/Blog.git">

  <meta name="octolytics-dimension-user_id" content="5421227" /><meta name="octolytics-dimension-user_login" content="petrkeil" /><meta name="octolytics-dimension-repository_id" content="32012442" /><meta name="octolytics-dimension-repository_nwo" content="petrkeil/Blog" /><meta name="octolytics-dimension-repository_public" content="true" /><meta name="octolytics-dimension-repository_is_fork" content="false" /><meta name="octolytics-dimension-repository_network_root_id" content="32012442" /><meta name="octolytics-dimension-repository_network_root_nwo" content="petrkeil/Blog" /><meta name="octolytics-dimension-repository_explore_github_marketplace_ci_cta_shown" content="false" />

    <link rel="canonical" href="https://github.com/petrkeil/Blog/blob/master/2015_03_12_R_ms_template/manuscript_template.Rmd" data-pjax-transient>

  <meta name="browser-stats-url" content="https://api.github.com/_private/browser/stats">

  <meta name="browser-errors-url" content="https://api.github.com/_private/browser/errors">

  <link rel="mask-icon" href="https://assets-cdn.github.com/pinned-octocat.svg" color="#000000">
  <link rel="icon" type="image/x-icon" class="js-site-favicon" href="https://assets-cdn.github.com/favicon.ico">

<meta name="theme-color" content="#1e2327">

  <link rel="manifest" href="/manifest.json" crossOrigin="use-credentials">

  </head>

  <body class="logged-in env-production page-blob">

  <div class="position-relative js-header-wrapper ">
    <a href="#start-of-content" tabindex="1" class="p-3 bg-blue text-white show-on-focus js-skip-to-content">Skip to content</a>
    <div id="js-pjax-loader-bar" class="pjax-loader-bar"><div class="progress"></div></div>

<header class="Header  f5" role="banner">
  <div class="d-flex flex-justify-between px-3 ">
    <div class="d-flex flex-justify-between ">
      <div class="">
        <a class="header-logo-invertocat" href="https://github.com/" data-hotkey="g d" aria-label="Homepage" data-ga-click="Header, go to dashboard, icon:logo">
  <svg height="32" class="octicon octicon-mark-github" viewBox="0 0 16 16" version="1.1" width="32" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0 0 16 8c0-4.42-3.58-8-8-8z"/></svg>
</a>

      </div>

    </div>

    <div class="HeaderMenu d-flex flex-justify-between flex-auto">
      <div class="d-flex">
            <div class="">
              <div class="header-search scoped-search site-scoped-search js-site-search position-relative js-jump-to"
  role="combobox"
  aria-owns="jump-to-results"
  aria-label="Search or jump to"
  aria-haspopup="listbox"
  aria-expanded="false"
>
  <div class="position-relative">
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-site-search-form" data-scope-type="Repository" data-scope-id="32012442" data-scoped-search-url="/petrkeil/Blog/search" data-unscoped-search-url="/search" action="/petrkeil/Blog/search" accept-charset="UTF-8" method="get"><input name="utf8" type="hidden" value="&#x2713;" />
      <label class="form-control header-search-wrapper header-search-wrapper-jump-to position-relative d-flex flex-justify-between flex-items-center js-chromeless-input-container">
        <input type="text"
          class="form-control header-search-input jump-to-field js-jump-to-field js-site-search-focus js-site-search-field is-clearable"
          data-hotkey="s,/"
          name="q"
          value=""
          placeholder="Search or jump to…"
          data-unscoped-placeholder="Search or jump to…"
          data-scoped-placeholder="Search or jump to…"
          autocapitalize="off"
          aria-autocomplete="list"
          aria-controls="jump-to-results"
          data-jump-to-suggestions-path="/_graphql/GetSuggestedNavigationDestinations#csrf-token=vKtepHuL5c+ELLxqvw5pBoFyrsH33qoqLeuwQXRDqVrb+0oabwda3WZ6kzvLvbGCX4m1kWaOg5yCDa6OI6VHtw=="
          spellcheck="false"
          autocomplete="off"
          >
          <input type="hidden" class="js-site-search-type-field" name="type" >
            <img src="https://assets-cdn.github.com/images/search-shortcut-hint.svg" alt="" class="mr-2 header-search-key-slash">

            <div class="Box position-absolute overflow-hidden d-none jump-to-suggestions js-jump-to-suggestions-container">
              <ul class="d-none js-jump-to-suggestions-template-container">
                <li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item" role="option">
                  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center p-2 jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open" href="">
                    <div class="jump-to-octicon js-jump-to-octicon mr-2 text-center d-none">
                      <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-octicon-repo d-none" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
                      <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-octicon-project d-none" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 0 0-1 1v14a1 1 0 0 0 1 1h13a1 1 0 0 0 1-1V1a1 1 0 0 0-1-1z"/></svg>
                      <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-octicon-search d-none" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0 0 13 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 0 0 0-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
                    </div>

                    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar d-none" alt="" aria-label="Team" src="" width="28" height="28">

                    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
                    </div>

                    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
                      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
                        In this repository
                      </span>
                      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
                        All GitHub
                      </span>
                      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
                    </div>

                    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
                      Jump to
                      <span class="d-inline-block ml-1 v-align-middle">↵</span>
                    </div>
                  </a>
                </li>
              </ul>
              <ul class="d-none js-jump-to-no-results-template-container">
                <li class="d-flex flex-justify-center flex-items-center p-3 f5 d-none">
                  <span class="text-gray">No suggested jump to results</span>
                </li>
              </ul>

              <ul id="jump-to-results" role="listbox" class="js-navigation-container jump-to-suggestions-results-container js-jump-to-suggestions-results-container" >
                <li class="d-flex flex-justify-center flex-items-center p-0 f5">
                  <img src="https://assets-cdn.github.com/images/spinners/octocat-spinner-128.gif" alt="Octocat Spinner Icon" class="m-2" width="28">
                </li>
              </ul>
            </div>
      </label>
</form>  </div>
</div>

            </div>

          <ul class="d-flex pl-2 flex-items-center text-bold list-style-none" role="navigation">
            <li>
              <a class="js-selected-navigation-item HeaderNavlink px-2" data-hotkey="g p" data-ga-click="Header, click, Nav menu - item:pulls context:user" aria-label="Pull requests you created" data-selected-links="/pulls /pulls/assigned /pulls/mentioned /pulls" href="/pulls">
                Pull requests
</a>            </li>
            <li>
              <a class="js-selected-navigation-item HeaderNavlink px-2" data-hotkey="g i" data-ga-click="Header, click, Nav menu - item:issues context:user" aria-label="Issues you created" data-selected-links="/issues /issues/assigned /issues/mentioned /issues" href="/issues">
                Issues
</a>            </li>
              <li>
                <a class="js-selected-navigation-item HeaderNavlink px-2" data-ga-click="Header, click, Nav menu - item:marketplace context:user" data-octo-click="marketplace_click" data-octo-dimensions="location:nav_bar" data-selected-links=" /marketplace" href="/marketplace">
                   Marketplace
</a>              </li>
            <li>
              <a class="js-selected-navigation-item HeaderNavlink px-2" data-ga-click="Header, click, Nav menu - item:explore" data-selected-links="/explore /trending /trending/developers /integrations /integrations/feature/code /integrations/feature/collaborate /integrations/feature/ship showcases showcases_search showcases_landing /explore" href="/explore">
                Explore
</a>            </li>
          </ul>
      </div>

      <div class="d-flex">

<ul class="user-nav d-flex flex-items-center list-style-none" id="user-links">
  <li class="dropdown">
    <span class="d-inline-block  px-2">

    <a aria-label="You have no unread notifications" class="notification-indicator tooltipped tooltipped-s  js-socket-channel js-notification-indicator" data-hotkey="g n" data-ga-click="Header, go to notifications, icon:read" data-channel="notification-changed:34245401" href="/notifications">
        <span class="mail-status "></span>
        <svg class="octicon octicon-bell" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M13.99 11.991v1H0v-1l.73-.58c.769-.769.809-2.547 1.189-4.416.77-3.767 4.077-4.996 4.077-4.996 0-.55.45-1 .999-1 .55 0 1 .45 1 1 0 0 3.387 1.229 4.156 4.996.38 1.879.42 3.657 1.19 4.417l.659.58h-.01zM6.995 15.99c1.11 0 1.999-.89 1.999-1.999H4.996c0 1.11.89 1.999 1.999 1.999z"/></svg>
</a>
    </span>
  </li>

  <li class="dropdown">
    <details class="details-overlay details-reset d-flex px-2 flex-items-center">
      <summary class="HeaderNavlink"
         aria-label="Create new…"
         data-ga-click="Header, create new, icon:add">
        <svg class="octicon octicon-plus float-left mr-1 mt-1" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 9H7v5H5V9H0V7h5V2h2v5h5v2z"/></svg>
        <span class="dropdown-caret mt-1"></span>
      </summary>
      <details-menu class="dropdown-menu dropdown-menu-sw">

<a role="menuitem" class="dropdown-item" href="/new" data-ga-click="Header, create new repository">
  New repository
</a>

  <a role="menuitem" class="dropdown-item" href="/new/import" data-ga-click="Header, import a repository">
    Import repository
  </a>

<a role="menuitem" class="dropdown-item" href="https://gist.github.com/" data-ga-click="Header, create new gist">
  New gist
</a>

  <a role="menuitem" class="dropdown-item" href="/organizations/new" data-ga-click="Header, create new organization">
    New organization
  </a>

  <div class="dropdown-divider"></div>
  <div class="dropdown-header">
    <span title="petrkeil/Blog">This repository</span>
  </div>
    <a role="menuitem" class="dropdown-item" href="/petrkeil/Blog/issues/new" data-ga-click="Header, create new issue">
      New issue
    </a>

      </details-menu>
    </details>
  </li>

  <li class="dropdown">

    <details class="details-overlay details-reset d-flex pl-2 flex-items-center">
      <summary class="HeaderNavlink name mt-1"
        aria-label="View profile and more"
        data-ga-click="Header, show menu, icon:avatar">
        <img alt="@jamesgrecian" class="avatar float-left mr-1" src="https://avatars2.githubusercontent.com/u/34245401?s=40&amp;v=4" height="20" width="20">
        <span class="dropdown-caret"></span>
      </summary>
      <details-menu class="dropdown-menu dropdown-menu-sw">
        <ul>
          <li class="header-nav-current-user css-truncate"><a role="menuitem" class="no-underline user-profile-link px-3 pt-2 pb-2 mb-n2 mt-n1 d-block" href="/jamesgrecian" data-ga-click="Header, go to profile, text:Signed in as">Signed in as <strong class="css-truncate-target">jamesgrecian</strong></a></li>
          <li class="dropdown-divider"></li>
          <li><a role="menuitem" class="dropdown-item" href="/jamesgrecian" data-ga-click="Header, go to profile, text:your profile">Your profile</a></li>
          <li><a role="menuitem" class="dropdown-item" href="/jamesgrecian?tab=repositories" data-ga-click="Header, go to repositories, text:your repositories">Your repositories</a></li>
          <li><a role="menuitem" class="dropdown-item" href="/jamesgrecian?tab=stars" data-ga-click="Header, go to starred repos, text:your stars">Your stars</a></li>
            <li><a role="menuitem" class="dropdown-item" href="https://gist.github.com/" data-ga-click="Header, your gists, text:your gists">Your gists</a></li>
          <li class="dropdown-divider"></li>
          <li><a role="menuitem" class="dropdown-item" href="https://help.github.com" data-ga-click="Header, go to help, text:help">Help</a></li>
          <li><a role="menuitem" class="dropdown-item" href="/settings/profile" data-ga-click="Header, go to settings, icon:settings">Settings</a></li>
          <li>
            <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="logout-form" action="/logout" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="70z+7pfJi/hSwFPV+HwBpRO+h6tPFRwPr5KWtHt/uVnEYJnva1K9ZZ8MF/preCNcmvA/GicmqsAz47+DjSrHhg==" />
              <button type="submit" class="dropdown-item dropdown-signout" data-ga-click="Header, sign out, icon:logout" role="menuitem">
                Sign out
              </button>
</form>          </li>
        </ul>
      </details-menu>
    </details>
  </li>
</ul>

        <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="sr-only right-0" action="/logout" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="xmQg0zHF9HnWv4MzKNWcgsKD5qm6fAiK0ZheRQvWYivtSEfSzV7C5Btzxxy70b57S81eGNJPvkVN6Xdy/YMc9A==" />
          <button type="submit" class="dropdown-item dropdown-signout" data-ga-click="Header, sign out, icon:logout">
            Sign out
          </button>
</form>      </div>
    </div>
  </div>
</header>

  </div>

  <div id="start-of-content" class="show-on-focus"></div>

    <div id="js-flash-container">

</div>

  <div role="main" class="application-main ">
        <div itemscope itemtype="http://schema.org/SoftwareSourceCode" class="">
    <div id="js-repo-pjax-container" data-pjax-container >

  <div class="pagehead repohead instapaper_ignore readability-menu experiment-repo-nav  ">
    <div class="repohead-details-container clearfix container">

      <ul class="pagehead-actions">
  <li>
        <!-- '"` --><!-- </textarea></xmp> --></option></form><form data-autosubmit="true" data-remote="true" class="js-social-container" action="/notifications/subscribe" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="oOjYFEFAFqKUnhayqPj0oBlERt/6I+ne5pf8Sl8iMXmjlzqHCag0+YA+/9+8U1yIKjTim6/xQpufP+1q6zhkGQ==" />      <input type="hidden" name="repository_id" id="repository_id" value="32012442" class="form-control" />

        <div class="select-menu js-menu-container js-select-menu">
          <a href="/petrkeil/Blog/subscription"
            class="btn btn-sm btn-with-count select-menu-button js-menu-target"
            role="button"
            aria-haspopup="true"
            aria-expanded="false"
            aria-label="Toggle repository notifications menu"
            data-ga-click="Repository, click Watch settings, action:blob#show">
            <span class="js-select-button">
                <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
                Watch
            </span>
          </a>
          <a class="social-count js-social-count"
            href="/petrkeil/Blog/watchers"
            aria-label="17 users are watching this repository">
            17
          </a>

        <div class="select-menu-modal-holder">
          <div class="select-menu-modal subscription-menu-modal js-menu-content">
            <div class="select-menu-header js-navigation-enable" tabindex="-1">
              <svg class="octicon octicon-x js-menu-close" role="img" aria-label="Close" viewBox="0 0 12 16" version="1.1" width="12" height="16"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
              <span class="select-menu-title">Notifications</span>
            </div>

              <div class="select-menu-list js-navigation-container" role="menu">

                <div class="select-menu-item js-navigation-item selected" role="menuitem" tabindex="0">
                  <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
                  <div class="select-menu-item-text">
                    <input type="radio" name="do" id="do_included" value="included" checked="checked" />
                    <span class="select-menu-item-heading">Not watching</span>
                    <span class="description">Be notified when participating or @mentioned.</span>
                    <span class="js-select-button-text hidden-select-button-text">
                      <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
                      Watch
                    </span>
                  </div>
                </div>

                <div class="select-menu-item js-navigation-item " role="menuitem" tabindex="0">
                  <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
                  <div class="select-menu-item-text">
                    <input type="radio" name="do" id="do_subscribed" value="subscribed" />
                    <span class="select-menu-item-heading">Watching</span>
                    <span class="description">Be notified of all conversations.</span>
                    <span class="js-select-button-text hidden-select-button-text">
                      <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
                        Unwatch
                    </span>
                  </div>
                </div>

                <div class="select-menu-item js-navigation-item " role="menuitem" tabindex="0">
                  <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
                  <div class="select-menu-item-text">
                    <input type="radio" name="do" id="do_ignore" value="ignore" />
                    <span class="select-menu-item-heading">Ignoring</span>
                    <span class="description">Never be notified.</span>
                    <span class="js-select-button-text hidden-select-button-text">
                      <svg class="octicon octicon-mute v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 2.81v10.38c0 .67-.81 1-1.28.53L3 10H1c-.55 0-1-.45-1-1V7c0-.55.45-1 1-1h2l3.72-3.72C7.19 1.81 8 2.14 8 2.81zm7.53 3.22l-1.06-1.06-1.97 1.97-1.97-1.97-1.06 1.06L11.44 8 9.47 9.97l1.06 1.06 1.97-1.97 1.97 1.97 1.06-1.06L13.56 8l1.97-1.97z"/></svg>
                        Stop ignoring
                    </span>
                  </div>
                </div>

              </div>

            </div>
          </div>
        </div>
</form>
  </li>

  <li>

  <div class="js-toggler-container js-social-container starring-container ">
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="starred js-social-form" action="/petrkeil/Blog/unstar" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="J1oHiD/jIasQWYvIghtqwNeIcq7Wz8HRwv5jjbAgOUiPVQ0EIQakzs+VFq2Jme9ff2CNRtGmfDd21a9Dmlzz+g==" />
      <input type="hidden" name="context" value="repository"></input>
      <button
        type="submit"
        class="btn btn-sm btn-with-count js-toggler-target"
        aria-label="Unstar this repository" title="Unstar petrkeil/Blog"
        data-ga-click="Repository, click unstar button, action:blob#show; text:Unstar">
        <svg class="octicon octicon-star v-align-text-bottom" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74L14 6z"/></svg>
        Unstar
      </button>
        <a class="social-count js-social-count" href="/petrkeil/Blog/stargazers"
           aria-label="79 users starred this repository">
          79
        </a>
</form>
    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="unstarred js-social-form" action="/petrkeil/Blog/star" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="QDgeOt/aBH8Yr862sLVj5rnSJEuVN8SGuxtswA5f/IDNGmAIWBfI4wwEyYlLxWpNLF8Shijy9vdBluRlDnHlzg==" />
      <input type="hidden" name="context" value="repository"></input>
      <button
        type="submit"
        class="btn btn-sm btn-with-count js-toggler-target"
        aria-label="Star this repository" title="Star petrkeil/Blog"
        data-ga-click="Repository, click star button, action:blob#show; text:Star">
        <svg class="octicon octicon-star v-align-text-bottom" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74L14 6z"/></svg>
        Star
      </button>
        <a class="social-count js-social-count" href="/petrkeil/Blog/stargazers"
           aria-label="79 users starred this repository">
          79
        </a>
</form>  </div>

  </li>

  <li>
          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="btn-with-count" action="/petrkeil/Blog/fork" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="Y84ilm5ix19gKu7BVYnfUlbhnPecmOc6pf0ITkJM8rS5L8RVxf9S1+VeyBggIdVBYUlUVnj+9FNoFw91y42Z/Q==" />
            <button
                type="submit"
                class="btn btn-sm btn-with-count"
                data-ga-click="Repository, show fork modal, action:blob#show; text:Fork"
                title="Fork your own copy of petrkeil/Blog to your account"
                aria-label="Fork your own copy of petrkeil/Blog to your account">
              <svg class="octicon octicon-repo-forked v-align-text-bottom" viewBox="0 0 10 16" version="1.1" width="10" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 1a1.993 1.993 0 0 0-1 3.72V6L5 8 3 6V4.72A1.993 1.993 0 0 0 2 1a1.993 1.993 0 0 0-1 3.72V6.5l3 3v1.78A1.993 1.993 0 0 0 5 15a1.993 1.993 0 0 0 1-3.72V9.5l3-3V4.72A1.993 1.993 0 0 0 8 1zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3 10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3-10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
              Fork
            </button>
</form>
    <a href="/petrkeil/Blog/network/members" class="social-count"
       aria-label="130 users forked this repository">
      130
    </a>
  </li>
</ul>

      <h1 class="public ">
  <svg class="octicon octicon-repo" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
  <span class="author" itemprop="author"><a class="url fn" rel="author" href="/petrkeil">petrkeil</a></span><!--
--><span class="path-divider">/</span><!--
--><strong itemprop="name"><a data-pjax="#js-repo-pjax-container" href="/petrkeil/Blog">Blog</a></strong>

</h1>

    </div>

<nav class="reponav js-repo-nav js-sidenav-container-pjax container"
     itemscope
     itemtype="http://schema.org/BreadcrumbList"
     role="navigation"
     data-pjax="#js-repo-pjax-container">

  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
    <a class="js-selected-navigation-item selected reponav-item" itemprop="url" data-hotkey="g c" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches repo_packages /petrkeil/Blog" href="/petrkeil/Blog">
      <svg class="octicon octicon-code" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M9.5 3L8 4.5 11.5 8 8 11.5 9.5 13 14 8 9.5 3zm-5 0L0 8l4.5 5L6 11.5 2.5 8 6 4.5 4.5 3z"/></svg>
      <span itemprop="name">Code</span>
      <meta itemprop="position" content="1">
</a>  </span>

    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
      <a itemprop="url" data-hotkey="g i" class="js-selected-navigation-item reponav-item" data-selected-links="repo_issues repo_labels repo_milestones /petrkeil/Blog/issues" href="/petrkeil/Blog/issues">
        <svg class="octicon octicon-issue-opened" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7 2.3c3.14 0 5.7 2.56 5.7 5.7s-2.56 5.7-5.7 5.7A5.71 5.71 0 0 1 1.3 8c0-3.14 2.56-5.7 5.7-5.7zM7 1C3.14 1 0 4.14 0 8s3.14 7 7 7 7-3.14 7-7-3.14-7-7-7zm1 3H6v5h2V4zm0 6H6v2h2v-2z"/></svg>
        <span itemprop="name">Issues</span>
        <span class="Counter">1</span>
        <meta itemprop="position" content="2">
</a>    </span>

  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
    <a data-hotkey="g p" itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_pulls checks /petrkeil/Blog/pulls" href="/petrkeil/Blog/pulls">
      <svg class="octicon octicon-git-pull-request" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 11.28V5c-.03-.78-.34-1.47-.94-2.06C9.46 2.35 8.78 2.03 8 2H7V0L4 3l3 3V4h1c.27.02.48.11.69.31.21.2.3.42.31.69v6.28A1.993 1.993 0 0 0 10 15a1.993 1.993 0 0 0 1-3.72zm-1 2.92c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zM4 3c0-1.11-.89-2-2-2a1.993 1.993 0 0 0-1 3.72v6.56A1.993 1.993 0 0 0 2 15a1.993 1.993 0 0 0 1-3.72V4.72c.59-.34 1-.98 1-1.72zm-.8 10c0 .66-.55 1.2-1.2 1.2-.65 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
      <span itemprop="name">Pull requests</span>
      <span class="Counter">0</span>
      <meta itemprop="position" content="3">
</a>  </span>

    <a data-hotkey="g b" class="js-selected-navigation-item reponav-item" data-selected-links="repo_projects new_repo_project repo_project /petrkeil/Blog/projects" href="/petrkeil/Blog/projects">
      <svg class="octicon octicon-project" viewBox="0 0 15 16" version="1.1" width="15" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 0 0-1 1v14a1 1 0 0 0 1 1h13a1 1 0 0 0 1-1V1a1 1 0 0 0-1-1z"/></svg>
      Projects
      <span class="Counter" >0</span>
</a>

    <a class="js-selected-navigation-item reponav-item" data-hotkey="g w" data-selected-links="repo_wiki /petrkeil/Blog/wiki" href="/petrkeil/Blog/wiki">
      <svg class="octicon octicon-book" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M3 5h4v1H3V5zm0 3h4V7H3v1zm0 2h4V9H3v1zm11-5h-4v1h4V5zm0 2h-4v1h4V7zm0 2h-4v1h4V9zm2-6v9c0 .55-.45 1-1 1H9.5l-1 1-1-1H2c-.55 0-1-.45-1-1V3c0-.55.45-1 1-1h5.5l1 1 1-1H15c.55 0 1 .45 1 1zm-8 .5L7.5 3H2v9h6V3.5zm7-.5H9.5l-.5.5V12h6V3z"/></svg>
      Wiki
</a>
  <a class="js-selected-navigation-item reponav-item" data-selected-links="repo_graphs repo_contributors dependency_graph pulse /petrkeil/Blog/pulse" href="/petrkeil/Blog/pulse">
    <svg class="octicon octicon-graph" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M16 14v1H0V0h1v14h15zM5 13H3V8h2v5zm4 0H7V3h2v10zm4 0h-2V6h2v7z"/></svg>
    Insights
</a>

</nav>

  </div>

<div class="container new-discussion-timeline experiment-repo-nav  ">
  <div class="repository-content ">

  <a class="d-none js-permalink-shortcut" data-hotkey="y" href="/petrkeil/Blog/blob/a32e8fd240fe0237ea9f64854aca0f7dfec0b7b6/2015_03_12_R_ms_template/manuscript_template.Rmd">Permalink</a>

  <!-- blob contrib key: blob_contributors:v21:0cba21dec00b9f9b447a2242a22fa2b0 -->

  <div class="file-navigation">

<div class="select-menu branch-select-menu js-menu-container js-select-menu float-left">
  <button class=" btn btn-sm select-menu-button js-menu-target css-truncate" data-hotkey="w"

    type="button" aria-label="Switch branches or tags" aria-expanded="false" aria-haspopup="true">
      <i>Branch:</i>
      <span class="js-select-button css-truncate-target">master</span>
  </button>

  <div class="select-menu-modal-holder js-menu-content js-navigation-container" data-pjax>

    <div class="select-menu-modal">
      <div class="select-menu-header">
        <svg class="octicon octicon-x js-menu-close" role="img" aria-label="Close" viewBox="0 0 12 16" version="1.1" width="12" height="16"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
        <span class="select-menu-title">Switch branches/tags</span>
      </div>

      <div class="select-menu-filters">
        <div class="select-menu-text-filter">
          <input type="text" aria-label="Filter branches/tags" id="context-commitish-filter-field" class="form-control js-filterable-field js-navigation-enable" placeholder="Filter branches/tags">
        </div>
        <div class="select-menu-tabs">
          <ul>
            <li class="select-menu-tab">
              <a href="#" data-tab-filter="branches" data-filter-placeholder="Filter branches/tags" class="js-select-menu-tab" role="tab">Branches</a>
            </li>
            <li class="select-menu-tab">
              <a href="#" data-tab-filter="tags" data-filter-placeholder="Find a tag…" class="js-select-menu-tab" role="tab">Tags</a>
            </li>
          </ul>
        </div>
      </div>

      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="branches" role="menu">

        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">

            <a class="select-menu-item js-navigation-item js-navigation-open selected"
               href="/petrkeil/Blog/blob/master/2015_03_12_R_ms_template/manuscript_template.Rmd"
               data-name="master"
               data-skip-pjax="true"
               rel="nofollow">
              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
              <span class="select-menu-item-text css-truncate-target js-select-menu-filter-text">
                master
              </span>
            </a>
        </div>

          <div class="select-menu-no-results">Nothing to show</div>
      </div>

      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="tags">
        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">

        </div>

        <div class="select-menu-no-results">Nothing to show</div>
      </div>

    </div>
  </div>
</div>

    <div class="BtnGroup float-right">
      <a href="/petrkeil/Blog/find/master"
            class="js-pjax-capture-input btn btn-sm BtnGroup-item"
            data-pjax
            data-hotkey="t">
        Find file
      </a>
      <clipboard-copy for="blob-path" class="btn btn-sm BtnGroup-item">
        Copy path
      </clipboard-copy>
    </div>
    <div id="blob-path" class="breadcrumb">
      <span class="repo-root js-repo-root"><span class="js-path-segment"><a data-pjax="true" href="/petrkeil/Blog"><span>Blog</span></a></span></span><span class="separator">/</span><span class="js-path-segment"><a data-pjax="true" href="/petrkeil/Blog/tree/master/2015_03_12_R_ms_template"><span>2015_03_12_R_ms_template</span></a></span><span class="separator">/</span><strong class="final-path">manuscript_template.Rmd</strong>
    </div>
  </div>

  <div class="commit-tease">
      <span class="float-right">
        <a class="commit-tease-sha" href="/petrkeil/Blog/commit/52b1e1cde0fb01bb0a205930ff265432581bb784" data-pjax>
          52b1e1c
        </a>
        <relative-time datetime="2015-03-12T22:12:59Z">Mar 12, 2015</relative-time>
      </span>
      <div>
        <img class="avatar" width="20" height="20" alt="" src="https://camo.githubusercontent.com/758da529902cc2f78e89a8a8ff28794146eabdf1/68747470733a2f2f322e67726176617461722e636f6d2f6176617461722f38626534333437643665363765386533353066616361316435613237663331643f643d68747470732533412532462532466173736574732d63646e2e6769746875622e636f6d253246696d6167657325324667726176617461727325324667726176617461722d757365722d3432302e706e6726723d6726733d313430" data-canonical-src="https://2.gravatar.com/avatar/8be4347d6e67e8e350faca1d5a27f31d?d=https%3A%2F%2Fassets-cdn.github.com%2Fimages%2Fgravatars%2Fgravatar-user-420.png&amp;r=g&amp;s=140" />
        <span class="user-mention">Petr Keil</span>
          <a data-pjax="true" title="Citations modified to include R and packages" class="message" href="/petrkeil/Blog/commit/52b1e1cde0fb01bb0a205930ff265432581bb784">Citations modified to include R and packages</a>
      </div>

    <div class="commit-tease-contributors">

<details class="details-reset details-overlay details-overlay-dark lh-default text-gray-dark float-left mr-2" id="blob_contributors_box">
  <summary class="btn-link" aria-haspopup="dialog"  >

    <span><strong>0</strong> contributors</span>
  </summary>
  <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast " aria-label="Users who have contributed to this file">
    <div class="Box-header">
      <button class="Box-btn-octicon btn-octicon float-right" type="button" aria-label="Close dialog" data-close-dialog>
        <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
      </button>
      <h3 class="Box-title">Users who have contributed to this file</h3>
    </div>

        <ul class="list-style-none overflow-auto">
        </ul>

  </details-dialog>
</details>

    </div>
  </div>

  <div class="file">
    <div class="file-header">
  <div class="file-actions">

    <div class="BtnGroup">
      <a id="raw-url" class="btn btn-sm BtnGroup-item" href="/petrkeil/Blog/raw/master/2015_03_12_R_ms_template/manuscript_template.Rmd">Raw</a>
        <a class="btn btn-sm js-update-url-with-hash BtnGroup-item" data-hotkey="b" href="/petrkeil/Blog/blame/master/2015_03_12_R_ms_template/manuscript_template.Rmd">Blame</a>
      <a rel="nofollow" class="btn btn-sm BtnGroup-item" href="/petrkeil/Blog/commits/master/2015_03_12_R_ms_template/manuscript_template.Rmd">History</a>
    </div>

          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="inline-form js-update-url-with-hash" action="/petrkeil/Blog/edit/master/2015_03_12_R_ms_template/manuscript_template.Rmd" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="RDQWb0ZLZztiOYIPlc8RR++SnYgTcU462l3VdiTyiYFsCe4KWd8HON+YIzeb3gW38TGuLa9sVdJc/AOC18Cn4w==" />
            <button class="btn-octicon tooltipped tooltipped-nw" type="submit"
              aria-label="Fork this project and edit the file" data-hotkey="e" data-disable-with>
              <svg class="octicon octicon-pencil" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M0 12v3h3l8-8-3-3-8 8zm3 2H1v-2h1v1h1v1zm10.3-9.3L12 6 9 3l1.3-1.3a.996.996 0 0 1 1.41 0l1.59 1.59c.39.39.39 1.02 0 1.41z"/></svg>
            </button>
</form>
        <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="inline-form" action="/petrkeil/Blog/delete/master/2015_03_12_R_ms_template/manuscript_template.Rmd" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="2SKvxp3pXN29EtFLN/3i3X0L0HbpmjB1H77ptWHdxXcWgHhUCtbdjiLn0Y5q/XAe46BnM5fLjZqhWOZPztrDLg==" />
          <button class="btn-octicon btn-octicon-danger tooltipped tooltipped-nw" type="submit"
            aria-label="Fork this project and delete the file" data-disable-with>
            <svg class="octicon octicon-trashcan" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 2H9c0-.55-.45-1-1-1H5c-.55 0-1 .45-1 1H2c-.55 0-1 .45-1 1v1c0 .55.45 1 1 1v9c0 .55.45 1 1 1h7c.55 0 1-.45 1-1V5c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm-1 12H3V5h1v8h1V5h1v8h1V5h1v8h1V5h1v9zm1-10H2V3h9v1z"/></svg>
          </button>
</form>  </div>

  <div class="file-info">
      112 lines (71 sloc)
      <span class="file-info-divider"></span>
    5.18 KB
  </div>
</div>

  <div id="readme" class="readme blob instapaper_body">
    <article class="markdown-body entry-content" itemprop="text"><table data-table-type="yaml-metadata">
  <thead>
  <tr>
  <th>title</th>
  <th>author</th>
  <th>date</th>
  <th>output</th>
  <th>csl</th>
  <th>bibliography</th>
  </tr>
  </thead>
  <tbody>
  <tr>
  <td><div>Template for writing scientific papers in R markdown</div></td>
  <td><div>Petr Keil, pkeil@seznam.cz</div></td>
  <td><div>11/1/2015</div></td>
  <td><div><table>
  <thead>
  <tr>
  <th>pdf_document</th>
  <th>html_document</th>
  </tr>
  </thead>
  <tbody>
  <tr>
  <td><div><table>
  <thead>
  <tr>
  <th>fig_caption</th>
  <th>keep_tex</th>
  <th>number_sections</th>
  </tr>
  </thead>
  <tbody>
  <tr>
  <td><div>true</div></td>
  <td><div>true</div></td>
  <td><div>true</div></td>
  </tr>
  </tbody>
</table>
</div></td>
  <td><div><table>
  <thead>
  <tr>
  <th>fig_caption</th>
  <th>force_captions</th>
  <th>highlight</th>
  <th>number_sections</th>
  <th>theme</th>
  </tr>
  </thead>
  <tbody>
  <tr>
  <td><div>true</div></td>
  <td><div>true</div></td>
  <td><div>pygments</div></td>
  <td><div>true</div></td>
  <td><div>cerulean</div></td>
  </tr>
  </tbody>
</table>
</div></td>
  </tr>
  </tbody>
</table>
</div></td>
  <td><div>mee.csl</div></td>
  <td><div>references.bib</div></td>
  </tr>
  </tbody>
</table>

<pre lang="{r,"><code>  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=FALSE)
  library(bibtex)
</code></pre>
<h1><a id="user-content-abstract" class="anchor" aria-hidden="true" href="#abstract"><svg class="octicon octicon-link" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Abstract</h1>
<p><em>Lorem ipsum dolor sit amet, est ad doctus eligendi scriptorem. Mel erat falli ut. Feugiat legendos adipisci vix at, usu at laoreet argumentum suscipiantur. An eos adhuc aliquip scriptorem, te adhuc dolor liberavisse sea. Ponderum vivendum te nec, id agam brute disputando mei.</em></p>
<h1><a id="user-content-introduction" class="anchor" aria-hidden="true" href="#introduction"><svg class="octicon octicon-link" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Introduction</h1>
<p>Lorem ipsum dolor sit amet, est ad doctus eligendi scriptorem. Mel erat falli ut. Feugiat legendos adipisci vix at, usu at laoreet argumentum suscipiantur. An eos adhuc aliquip scriptorem, te adhuc dolor liberavisse sea. Ponderum vivendum te nec, id agam brute disputando mei.</p>
<p>Putant numquam tacimates at eum. Aliquip torquatos ex vis, mei et quando debitis appareat, impetus accumsan corrumpit in usu. Nam mucius facilis singulis id, duo ei autem imperdiet instructior. Cu ceteros alienum mel, id vix putant impedit, ex idque eruditi forensibus eum. Posse dicunt id usu. Ei iracundia constituto sed, duo ne exerci ignota, an eum unum conceptam.</p>
<p>Has audire salutandi no, ut eam dicat libris dicunt. Pri hendrerit quaerendum adversarium ea, dicat atqui munere et sea. Illum insolens eos ne, eu enim graece rationibus mea. At postea utamur mel, eius nonumes percipitur at vis. Numquam similique in per, te quo saepe utroque pericula.</p>
<p>Ea nonumy volumus usu, no mel inermis dissentias. Dico partiendo vituperatoribus eum et. Mea accusam convenire te, usu populo qualisque gloriatur ut. Eu eum oratio altera option, ad mea ignota scriptorem. Ne suas latine vix, eos oblique sanctus pertinax cu.</p>
<h1><a id="user-content-methods" class="anchor" aria-hidden="true" href="#methods"><svg class="octicon octicon-link" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Methods</h1>
<p>Lorem ipsum dolor sit amet, est ad doctus eligendi scriptorem. Mel erat falli ut. Feugiat legendos adipisci vix at, usu at laoreet argumentum suscipiantur. An eos adhuc aliquip scriptorem, te adhuc dolor liberavisse sea. Ponderum vivendum te nec, id agam brute disputando mei.</p>
<p>Putant numquam tacimates at eum. Aliquip torquatos ex vis, mei et quando debitis appareat, impetus accumsan corrumpit in usu. Nam mucius facilis singulis id, duo ei autem imperdiet instructior. Cu ceteros alienum mel, id vix putant impedit, ex idque eruditi forensibus eum. Posse dicunt id usu. Ei iracundia constituto sed, duo ne exerci ignota, an eum unum conceptam.</p>
<h2><a id="user-content-equations" class="anchor" aria-hidden="true" href="#equations"><svg class="octicon octicon-link" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Equations</h2>
<p>The deterministic part of the model is defined by this <strong>in-line equation</strong> as
$\mu_i = \beta_0 + \beta_1x$, and the stochastic part by the <strong>centered equation</strong>:</p>
<p>$$ \frac{1}{\sqrt{2\pi}\sigma}e^{-(x-\mu_i)^2/(2\sigma^2)} $$</p>
<h2><a id="user-content-tables" class="anchor" aria-hidden="true" href="#tables"><svg class="octicon octicon-link" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Tables</h2>
<pre lang="{r"><code>library(knitr)

x &lt;- rnorm(100)
y &lt;- 2*x + rnorm(100)
out &lt;- lm(y ~ x)
kable(summary(out)$coef, digits=2, caption="This is a GLM summary table.")
</code></pre>
<pre lang="{r"><code>library(xtable)
tab &lt;- xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
              caption="This is a GLM summary table.")
print(tab, type="latex")
</code></pre>
<h2><a id="user-content-plots" class="anchor" aria-hidden="true" href="#plots"><svg class="octicon octicon-link" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Plots</h2>
<pre lang="{r"><code>  plot(x, y)
  abline(out)
</code></pre>
<h2><a id="user-content-citations" class="anchor" aria-hidden="true" href="#citations"><svg class="octicon octicon-link" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Citations</h2>
<p>The relationship was first described by
<code>r citet("10.1111/j.1461-0248.2005.00827.x")</code>. However, there are also opinions
that the relationship is spurious <code>r citep("10.1111/j.2041-210x.2012.00264.x")</code>.
We used R for our calculations <code>r citep(citation())</code>, and we used package <code>knitcitations</code> <code>r citep(citation("knitcitations"))</code> to make the bibliography.</p>
<h1><a id="user-content-results-and-discussion" class="anchor" aria-hidden="true" href="#results-and-discussion"><svg class="octicon octicon-link" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Results and discussion</h1>
<p>Lorem ipsum dolor sit amet, est ad doctus eligendi scriptorem. Mel erat falli ut. Feugiat legendos adipisci vix at, usu at laoreet argumentum suscipiantur. An eos adhuc aliquip scriptorem, te adhuc dolor liberavisse sea. Ponderum vivendum te nec, id agam brute disputando mei.</p>
<p>Putant numquam tacimates at eum. Aliquip torquatos ex vis, mei et quando debitis appareat, impetus accumsan corrumpit in usu. Nam mucius facilis singulis id, duo ei autem imperdiet instructior. Cu ceteros alienum mel, id vix putant impedit, ex idque eruditi forensibus eum. Posse dicunt id usu. Ei iracundia constituto sed, duo ne exerci ignota, an eum unum conceptam.</p>
<p>Has audire salutandi no, ut eam dicat libris dicunt. Pri hendrerit quaerendum adversarium ea, dicat atqui munere et sea. Illum insolens eos ne, eu enim graece rationibus mea. At postea utamur mel, eius nonumes percipitur at vis. Numquam similique in per, te quo saepe utroque pericula.</p>
<p>Ea nonumy volumus usu, no mel inermis dissentias. Dico partiendo vituperatoribus eum et. Mea accusam convenire te, usu populo qualisque gloriatur ut. Eu eum oratio altera option, ad mea ignota scriptorem. Ne suas latine vix, eos oblique sanctus pertinax cu.</p>
<h1><a id="user-content-references" class="anchor" aria-hidden="true" href="#references"><svg class="octicon octicon-link" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>References</h1>
<pre lang="{r,"><code>  write.bibtex(file="references.bib")
</code></pre>
</article>
  </div>

  </div>

  <details class="details-reset details-overlay details-overlay-dark">
    <summary data-hotkey="l" aria-label="Jump to line"></summary>
    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast linejump" aria-label="Jump to line">
      <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-jump-to-line-form Box-body d-flex" action="" accept-charset="UTF-8" method="get"><input name="utf8" type="hidden" value="&#x2713;" />
        <input class="form-control flex-auto mr-3 linejump-input js-jump-to-line-field" type="text" placeholder="Jump to line&hellip;" aria-label="Jump to line" autofocus>
        <button type="submit" class="btn" data-close-dialog>Go</button>
</form>    </details-dialog>
  </details>

  </div>
  <div class="modal-backdrop js-touch-events"></div>
</div>

    </div>
  </div>

  </div>

<div class="footer container-lg px-3" role="contentinfo">
  <div class="position-relative d-flex flex-justify-between pt-6 pb-2 mt-6 f6 text-gray border-top border-gray-light ">
    <ul class="list-style-none d-flex flex-wrap ">
      <li class="mr-3">&copy; 2018 <span title="0.17379s from unicorn-6bfd5c759c-j6nwh">GitHub</span>, Inc.</li>
        <li class="mr-3"><a data-ga-click="Footer, go to terms, text:terms" href="https://github.com/site/terms">Terms</a></li>
        <li class="mr-3"><a data-ga-click="Footer, go to privacy, text:privacy" href="https://github.com/site/privacy">Privacy</a></li>
        <li class="mr-3"><a href="https://help.github.com/articles/github-security/" data-ga-click="Footer, go to security, text:security">Security</a></li>
        <li class="mr-3"><a href="https://status.github.com/" data-ga-click="Footer, go to status, text:status">Status</a></li>
        <li><a data-ga-click="Footer, go to help, text:help" href="https://help.github.com">Help</a></li>
    </ul>

    <a aria-label="Homepage" title="GitHub" class="footer-octicon mr-lg-4" href="https://github.com">
      <svg height="24" class="octicon octicon-mark-github" viewBox="0 0 16 16" version="1.1" width="24" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0 0 16 8c0-4.42-3.58-8-8-8z"/></svg>
</a>
   <ul class="list-style-none d-flex flex-wrap ">
        <li class="mr-3"><a data-ga-click="Footer, go to contact, text:contact" href="https://github.com/contact">Contact GitHub</a></li>
        <li class="mr-3"><a href="https://github.com/pricing" data-ga-click="Footer, go to Pricing, text:Pricing">Pricing</a></li>
      <li class="mr-3"><a href="https://developer.github.com" data-ga-click="Footer, go to api, text:api">API</a></li>
      <li class="mr-3"><a href="https://training.github.com" data-ga-click="Footer, go to training, text:training">Training</a></li>
        <li class="mr-3"><a href="https://blog.github.com" data-ga-click="Footer, go to blog, text:blog">Blog</a></li>
        <li><a data-ga-click="Footer, go to about, text:about" href="https://github.com/about">About</a></li>

    </ul>
  </div>
  <div class="d-flex flex-justify-center pb-6">
    <span class="f6 text-gray-light"></span>
  </div>
</div>

  <div id="ajax-error-message" class="ajax-error-message flash flash-error">
    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 0 0 0 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 0 0 .01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
    <button type="button" class="flash-close js-ajax-error-dismiss" aria-label="Dismiss error">
      <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
    </button>
    You can’t perform that action at this time.
  </div>

    <script crossorigin="anonymous" integrity="sha512-RJ1ufbxsSbKjRCyYvinsPNKvTBvcvvKUvEOJ8g+GjtWI5SuRr+QPBlZcvRDws4H9YwAgdQFcGxfL8UbwEfdI7A==" type="application/javascript" src="https://assets-cdn.github.com/assets/compat-daf14de28fadf1e2bc40d100cb773e2b.js"></script>
    <script crossorigin="anonymous" integrity="sha512-IC5Ai6c7cUBfMx54JUAsIjA7VroipzkizlZ8F6/0efsGGyZStweZV7bX9bxSFL07nUNLfwrKSEysVIijX+CeUw==" type="application/javascript" src="https://assets-cdn.github.com/assets/frameworks-5874e38105ec062c740350eeb96bdd8a.js"></script>

    <script crossorigin="anonymous" async="async" integrity="sha512-/E7YUPwg5s5FmwhSiwtXwNsTUCpjjdcrKoK7UFMQCguTSC71dACKl+XDhBftzkBsRra4w+a24G4oU535TieAIA==" type="application/javascript" src="https://assets-cdn.github.com/assets/github-1835546276a57f4d0c3203810d2ec7d6.js"></script>

  <div class="js-stale-session-flash stale-session-flash flash flash-warn flash-banner d-none">
    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 0 0 0 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 0 0 .01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
    <span class="signed-in-tab-flash">You signed in with another tab or window. <a href="">Reload</a> to refresh your session.</span>
    <span class="signed-out-tab-flash">You signed out in another tab or window. <a href="">Reload</a> to refresh your session.</span>
  </div>
  <div class="facebox" id="facebox" style="display:none;">
  <div class="facebox-popup">
    <div class="facebox-content" role="dialog" aria-labelledby="facebox-header" aria-describedby="facebox-description">
    </div>
    <button type="button" class="facebox-close js-facebox-close" aria-label="Close modal">
      <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
    </button>
  </div>
</div>

  <template id="site-details-dialog">
  <details class="details-reset details-overlay details-overlay-dark lh-default text-gray-dark" open>
    <summary aria-haspopup="dialog" aria-label="Close dialog"></summary>
    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast">
      <button class="Box-btn-octicon m-0 btn-octicon position-absolute right-0 top-0" type="button" aria-label="Close dialog" data-close-dialog>
        <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
      </button>
      <div class="octocat-spinner my-6 js-details-dialog-spinner"></div>
    </details-dialog>
  </details>
</template>

  <div class="Popover js-hovercard-content position-absolute" style="display: none; outline: none;" tabindex="0">
  <div class="Popover-message Popover-message--bottom-left Popover-message--large Box box-shadow-large" style="width:360px;">
  </div>
</div>

<div id="hovercard-aria-description" class="sr-only">
  Press h to open a hovercard with more details.
</div>

  </body>
</html>