### 2018-10-03 manuscript/manuscript_template.Rmd
Update manuscript to include outline abstract

---
title:  'Incorporating dive behaviour into models of marine animal movement'

author: |
  | W. J. Grecian$^1$ ^[Corresponding author: james.grecian@gmail.com] , I. D. Jonsen$^2$
  | $^1$Sea Mammal Research Unit, Scottish Oceans Institute, University of St Andrews
  | $^2$Dept. of Biological Sciences, Macquarie University, Sydney, Australia

abstract: |
 The Abstract must not exceed 350 words and should list the main results and conclusions, using simple, factual, numbered statements.  
 1. Models of animal movement often only consider horizontal movement over 2 dimensions. Incorporating vertical movement in these models offers the potential to improve our understanding of the drivers of animal movement.  
 2. Here we describe a state-space model that incorporates a random walk on both horizontal and vertical movements, and estimates the covariance between them. The model is flexible to allow both parameters to be conditioned on covariates, offering insights into the drivers of animal movement.  
 3. outline the main results  
 4. identify the conclusions and the wider implications.

keywords: | 
 A list in alphabetical order not exceeding eight words or short phrases. The most important key-words should appear in the title and the abstract as well as the key-word list.

date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: cerulean
csl: mee.csl
#bibliography: references.bib
header-includes:
   - \usepackage{lineno}
   - \linenumbers
---

```{r, echo=FALSE}
  # devtools::install_github("cboettig/knitcitations@v1")
  library(knitcitations); cleanbib()
  cite_options(citation_format = "pandoc", check.entries=TRUE)
  library(bibtex)
```

# INTRODUCTION

Lorem ipsum dolor sit amet, est ad doctus eligendi scriptorem. Mel erat falli ut. Feugiat legendos adipisci vix at, usu at laoreet argumentum suscipiantur. An eos adhuc aliquip scriptorem, te adhuc dolor liberavisse sea. Ponderum vivendum te nec, id agam brute disputando mei.

# MATERIALS AND METHODS

Lorem ipsum dolor sit amet, est ad doctus eligendi scriptorem. Mel erat falli ut. Feugiat legendos adipisci vix at, usu at laoreet argumentum suscipiantur. An eos adhuc aliquip scriptorem, te adhuc dolor liberavisse sea. Ponderum vivendum te nec, id agam brute disputando mei.

## Equations

The deterministic part of the model is defined by this **in-line equation** as 
$\mu_i = \beta_0 + \beta_1x$, and the stochastic part by the **centered equation**:

$$ \frac{1}{\sqrt{2\pi}\sigma}e^{-(x-\mu_i)^2/(2\sigma^2)} $$

## Tables

```{r kable, echo=FALSE}
library(knitr)

x <- rnorm(100)
y <- 2*x + rnorm(100)
out <- lm(y ~ x)
kable(summary(out)$coef, digits=2, caption="This is a GLM summary table.")
```

```{r xtable, results="asis", echo=FALSE, eval=FALSE, message=FALSE}
library(xtable)
tab <- xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
              caption="This is a GLM summary table.")
print(tab, type="latex")
```

## Plots

```{r carDataPlot, echo=FALSE, fig.cap="Relationship between x and y. The solid line is least-squares linear regression.", fig.width=4, fig.height=4}
  plot(x, y)
  abline(out)
```

## Citations

The relationship was first described by @Halpern_2006. However, there are also opinions
that the relationship is spurious [@Keil_2012].
We used R for our calculations [@R_Core_Team_2016], and we used package `knitcitations` [@Boettiger_2017] to make the bibliography.

# RESULTS

Lorem ipsum dolor sit amet, est ad doctus eligendi scriptorem. Mel erat falli ut. Feugiat legendos adipisci vix at, usu at laoreet argumentum suscipiantur. An eos adhuc aliquip scriptorem, te adhuc dolor liberavisse sea. Ponderum vivendum te nec, id agam brute disputando mei.

# DISCUSSION

# CONCLUSION

# ACKNOWLEDGEMENTS

This project was made possible through a travel award from the Canada-UK Foundation.

# AUTHORS' CONTRIBUTION

# DATA ACCESSIBILITY

# REFERENCES

```{r, warning=FALSE, message=FALSE, echo=FALSE} 
  #write.bibtex(file="references.bib")
```